As I reflect on my journey at Princeton, I am forever grateful for those who pushed me to grow, kept me grounded, and made sure I never forgot to laugh along the way.

To Professor Akrotirianakis. Thank you for your mentorship and for never once making me feel like this was a bad idea. To Stefan and the rest of my research group. Thank you for watching this thesis crawl from a half-baked idea into something real.

To Mr. Rucker, my Jefe. Thank you for igniting my passion for learning. There was never a boring day in your class---and I still don't know how you tolerated the barrage of questions I deliberately saved until the bell rang. \#Jefites4Life

To Maxwel. Thank you for being unapologetically yourself. There's never a dull moment around you. So---PPL and Small World at 11?

To Ethan and Alex, my fellow ORFE classmates. I can confidently say at least one of us knew what was going on at any given time. Shoutout to whoever that was.

To the badminton team: thank you for every Thursday night. My only regret is not exploiting you guys more for late meals.

To my brothers in PBJ. The things we've lost at that table---money, sleep, my faith in the law of large numbers. Still, I wouldn't trade a single wee hour for anything.

To The Goonies. Thank you for everything. The memories, the laughs, and the poffertjes---which, for the record, we discovered in Amsterdam and have absolutely nothing to do with those Dutch mini pancakes.

To everyone in Bree Flowies. You've known me since before any of this, which means you have the most dirt on me. Thank you for your continued silence.

And last but not least, to my family. None of this happens without you. Thank you for the love, the sacrifices, the unwavering belief in me, and for always reminding me that home is wherever you are.

(This thesis was copy-edited and proofread with assistance from AI tools, including Claude, ChatGPT, and Gemini. They also assisted with code generation.)